---
title: "Analysis Script"
output: html_notebook
---

```{r}
library(tidyverse)
source("cleaning_script.R")
```
Some more exploratory analysis first:
```{r}
brisbane_complaints %>% 
  group_by(type_of_animal) %>% 
  ggplot(aes(x = type_of_animal, col = complaint, fill = complaint)) +
  geom_bar(position = "fill")

brisbane_complaints %>% 
  group_by(type_of_animal) %>% 
  ggplot(aes(x = type_of_animal, fill = suburb)) +
  geom_bar()
```

```{r}
animal_complaints %>% 
  group_by(complaint_type) %>% 
  ggplot(aes(x = complaint_type, fill = animal_type)) +
  geom_bar(position = "fill")

animal_complaints %>% 
  group_by(date_received) %>% 
  summarise(count = n()) %>% 
  ggplot(aes(x = date_received, y = count)) +
  geom_line()
```

What are the challenges I am facing now? 

1) There are far too many suburbs in the Brisbane dataset. How do I create any useful visualisation or analysis with so much areas?

2) I need more clarity and to be more specific in what I am doing. I am currently still in "explore" mode, but I may need to pick a question and focus on that. As it stands, it feels like I'm making no progress and running around in circles, I need to stay on track and focus. Keeping the questions in mind at all times is very important for this.








